---
title: "STAT 585 Lab One"
author: "Eryn Blagg, Katherine Goode, Ju Ji, Charlotte Roiger"
date: "2/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, echo = FALSE}
# Load in requisite libraries
# install.packages(c("maps", "mapdata"))

library(tidyverse)
library(ggmap)
library(maps)
library(mapdata)

```

# Part One: Building the Book

**Deliverable (1):** The last word of Chapter 15 is “packages”

**Deliverable (2):** The biggest problem we encountered when trying to build the book is to install all of the packages that the book required to compile. Everytime we knitted the book, the execution was halted and `R` would indicate which package would need to be installed. The following packages were required for installation: sloop, vctrs, RSQLite, profvis, bench, ggbeeswarm. Another important thing to note is that the book required the package “emo” which was not available on CRAN. Because it was not available on CRAN we had to use the `devtools` to install “emo” from Hadley’s Git Repo.

# Part Two: US Weather Stations


```{r, echo=FALSE, message = FALSE}
# Read in the data for each directory

# Make sure your working directory is pointed to the right file path for this lab please
dataDir <- getwd()

# Load in data, chosen to specify start and stopping point for each column in data frame
weather1 <- read_fwf(file.path(dataDir,"ushcn-v2.5-stations.txt"), fwf_cols(country_code = c(1,2), network_code = 3, id_placeholders = c(4,5), coop_id = c(6,11), latitude = c(13,20), longitude = c(22,30), elevation = c(33,37), state = c(39,40), name = c(42,71), comp1 = c(73,78), comp2 = c(80,85), comp3 = c(87,92), utc_offset = c(94,95)))

# Will also use maps from the mapdata package for visualization
usa <- map_data("state")

```


```{r echo = FALSE}
# Change levation to be a numeric rather than factor
weather1 <- weather1 %>%
  mutate(elevation = as.numeric(elevation))

# Create a variable that accounts for the timezone of each state 
# THERE HAS TO BE A BETTER WAY TO DO THIS
usa <- usa %>%
  mutate(time_zone = ifelse(region %in% c("maine", "vermont", "new hampshire", "massachusetts", "rhode island", "conneticut", "new york", "pennsylvania", "new jersey", "delaware", "maryland", "district of columbia", "virginia", "west virginia", "ohio", "michigan", "indiana", "north carolina", "south carolina", "georgia", "florida", "connecticut"), "ETZ",
                            ifelse(region %in% c("wisconsin", "illinois", "kentucky", "tennessee", "alabama", "mississippi", "minnesota", "iowa", "missouri", "arkansas", "louisiana", "north dakota", "south dakota", "nebraska", "kansas", "oklahoma", "texas", ""), "CTZ", 
                                   ifelse( region %in% c("montana", "wyoming", "colorado", "new mexico", "idaho", "utah", "arizona"), "MTZ", "PTZ"))))

  
```


```{r}
# Visualize elevation, state, and time zone of each station

ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group, fill = time_zone, alpha = 0.2), color = "black") + 
  geom_point(data = weather1, aes(x = longitude, y = latitude, colour = elevation))+
  coord_fixed(1.3)
```

# More Weather Stations

```{r}
# ERROR 

# fn <- "ftp://ftp.ncdc.noaa.gov/pub/data/ushcn/v2.5/ushcn.tavg.latest.raw.tar.gz"
# download.file(fn,destfile="tmp.tar.gz")
# untar("tmp.tar.gz",files="tavg.latest.csv")
# X <- read.csv("tavg.latest.csv")
```

